### 2025-08-29 sections/4-results.tex
added balanced vs unbalanced to results

@@

>>>>

\subsection{Scaling Behavior and Recall}
\label{sec:results:scaling-behavior-and-recall}

@@

\subsection{Clustering Strategies and Number of Distance Computations}
\label{sec:results:clustering-strategies-and-number-of-distance-computations}

In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
These strategies are the Cartesian product of balanced vs. unbalanced clustering, and the presence vs. absence of depth-first reordering as described in Section~\ref{sec:methods:clustering:depth-first-reordering}.
To help with this analysis, we also added instrumentation to our implementation of CAKES to count the number of distance computations during search.
We note that this instrumentation significantly slows down the wall-clock time of search, so it would not be used in a real-world application.

Figure~\ref{fig:results:distance-counts}. shows these four comparisons (referred to as Ball, BalancedBall, PermutedBall, and, PermutedBalancedBall, where depth-first-reordering is called Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset.
Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse;
the algorithms (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots} ( which shows results using PermutedBall), no longer do so under a balanced clustering.
We note that on this dataset, depth-first reordering (the Permuted variants) seems to have little effect on throughput; depth-first reordering is primarily intended to improve space complexity.

\begin{figure}[h]
  \centering
  \subfloat[Repeated $\rho$-NN]{
  \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_throughput.png}
  \label{fig:results:fashion-mnist-rnn-throughput}
  }
  \subfloat[Repeated $\rho$-NN]{
    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnRepeatedRnn_10_counts.png}    
    \label{fig:results:fashion-mnist-rnn-counts}
  }

  \subfloat[Breadth First Sieve]{
    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_throughput.png}
    \label{fig:results:fashionmnist-bfs-throughput}
    }
  \subfloat[Breadth First Sieve]{
    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnBreadthFirst_10_counts.png}
    \label{fig:results:fashionmnist-bfs-counts}
    }

  \subfloat[Depth First Sieve]{
    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_throughput.png}
    \label{fig:results:fashionmnist-dfs-throughput}
    }
  \subfloat[Depth First Sieve]{
    \includegraphics[width=0.48\columnwidth]{images/distance_counts/fashion-mnist_KnnDepthFirst_10_counts.png}
    \label{fig:results:fashionmnist-dfs-counts}
  }

  \vspace{2pt}
  \includegraphics[width=1\columnwidth]{images/distance_counts/legend.png}

  \caption{Number of distance computations across four clustering strategies and three search algorithms on the Fashion-MNIST dataset.
  Adding the instrumentation to count the number of distance computations had the side-effect of significantly slowing down the search algorithms compared to those reported in Figure~\ref{fig:results:scaling-plots}.
  The left column shows the throughput in queries per second, while the right column shows the mean number of distance computations per query.
  The x-axis represents increasing cardinality of the dataset.}
  \label{fig:results:distance-counts}
\end{figure}
